---
title: "Process_mediation"
author: "R. Gaizauskaite"
format: html
editor: visual
---

## Library

```{r}
library(tidyverse)
```

## Data

```{r}
CST_data <- read_csv("imputed_complete_data.csv")
```

## Running mediation (simple)

```{r}
process (data = CST_data, y = "CST", x = "Sex", m ="EA2", model = 4)
```

Indirect effect is significant, however direct effect stays significant also, so indirect effect do not explain all the difference.

```{r}
process (data = CST_data, y = "CST", x = "Sex", m ="EA2", model = 4, effsize =1, total =1, stand =1, boot = 10000 , modelbt = 1, seed = 654321)
```

Procentage of explain effect EA2

```{r}
Proc_EA <- (2.1678*100)/13.2595
Proc_EA
```

Negative emotional arousal ir data explains about 16 of the difference between sex'es.

## With STEM as co-variate

**Final model**

```{r}
process (data = CST_data, y = "CST", x = "Sex", m ="EA2", model = 4, effsize =1, total =1, stand =1, cov = "STEM", boot = 10000 , modelbt = 1, seed = 654321)
```

```{}
```

# With missing values

Uploading dataset

```{r}
CST_missing <- read_csv("data_final_plus.csv") |> filter(Subject != "128") |> drop_na(Emo_arousal_afterER)
```

## Simple mediation

```{r}
process (data = CST_missing, y = "CorrectA", x = "Lyt", m ="Emo_arousal_afterER", model = 4)
```

With missing values results the same, indirect effect is significant.

```{r}
process (data = CST_missing, y = "CorrectA", x = "Lyt", m ="Emo_arousal_afterER", model = 4, effsize =1, total =1, stand =1, cov = "STEM", boot = 10000 , modelbt = 1, seed = 654321)
```

All good, results basically are the same, therefore we can use imputed data, as it has higher power.
