---
title: Main analysis with missing values
author: R. Gaizauskaite
format: 
  html:
    toc: true
    embed-resources: true
---

### Libraries

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(rstatix)
library(gtsummary)
library(lmtest)
library(ggpubr)
library(car)
library(MASS)
library(nortest)
```

```{r}
#| echo: false
combined_full <- read_csv("C:/Users/Rokas/Desktop/Mantes dokumentai/Dok. reikalingi failai/Skerspjūvio analizė/final_final_analysis/Full_CST_analysis/Data_with_missing.csv")
```

```{r}
#| echo: false
CST_full_parameters <- combined_full |> mutate(accu = (Orthogon_CorrectSimple + Orthogon_CorrectJoint + Orthogon_CorrectEmbed + Oblique_CorrectSimple + Oblique_CorrectJoint + Oblique_CorrectEmbed)/6)
```

```{r}
#| echo: false
CST_full_parameters <- CST_full_parameters |>  
                          mutate(Group = as.factor(Group),
                                STEM = as.factor(STEM), 
                                Sex =  as.factor(Sex))
```

```{r}
#| echo: false
# Define the desired order of factor levels
desired_order <- c("Men", "NCF", "NCL", "OC", "IUD" )

# Reorder the levels of the 'Group' factor variable
CST_full_parameters$Group <- fct_relevel(CST_full_parameters$Group, desired_order)

```

```{r}
#| echo: false
CST_full_parameters <- CST_full_parameters |> 
  mutate(Group = fct_recode(Group, "Males" = "Men"))
```

## 1. Main analysis

### 1. 1 Demographic table

```{r}
Demographic_data <- CST_full_parameters |> dplyr::select(Age:Tes,-Sex, -anxiety)

demographic_summary_1 <- Demographic_data |>
  tbl_summary(
    by = Group,
    missing = "no",
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    )
  ) |>
  add_p(
    test = all_continuous() ~ "aov",
    pvalue_fun = function(x) style_pvalue(x, digits = 2)
  )

demographic_summary_1
```



### 1.3. Linear regression model

### 1.3.1 Women

```{r}
#| echo: false
women_data <- CST_full_parameters |> filter(Sex == 1)
```

```{r}
#| echo: false
dummy_matrix <- model.matrix(~ Group - 1, data = women_data)
dummy_df <- as.data.frame(dummy_matrix)

#Combining the dummy variables with the original data
women_data <- cbind(women_data, dummy_df)


women_data <- women_data[, -26] 
```

```{r}
#| echo: false
women_data <- women_data |> rename(
  OC = GroupOC,
  IUD = GroupIUD,
  NCF = GroupNCF,
  NCL = GroupNCL
  )
```

```{r}
#| echo: false
women_data <- women_data |> 
  mutate(
    OC = as.factor(OC),
    IUD = as.factor(IUD),
    NCF = as.factor(NCF),
    NCL = as.factor(NCL)
  )
```

```{r}
women_data <- women_data |> drop_na( EA1, EA2, Tiredness_1, Tiredness_2, P4, E2, Tes)
```


#### Running first LM model with all the variables

```{r}
women_lm_model <- lm(accu~ OC + IUD + NCL + Age + STEM + PANAS_PA + PANAS_NA + EA1 + EA2 + Tiredness_1 + Tiredness_2 + Femininity + Masculinity + P4 + E2 +Tes, data = women_data)
```

```{r}
summary(women_lm_model)
```

#### Multicolinearity check up

```{r}
vif(women_lm_model)
```

All VIF score \< 3, no multicolinearity

#### AIC procedure

```{r}
stepAIC(women_lm_model, direction = "backward")
```

As model suggest IUD group to stay, i would do little bit differently and IUD group leave as reference group to account for all hormonal status groups.

#### Final model

```{r}
women_final <- lm(accu ~ OC + NCF + NCL + EA2 + Femininity + STEM, data  = women_data)
```

```{r}
summary(women_final)
```

#### Assumption testing

```{r}
#| echo: false
par(mfrow = c(2, 2))
plot(women_final)
```

**Normality of residuals**

```{r}
#| echo: false
#| output: false
model.metrics.women <- augment(women_final) |> 
  dplyr::select(-.hat, -.sigma, -.fitted)

shapiro_test(model.metrics.women$.resid)
```

Normality of residual assuption is not violated, p \> 0.05

**Homoscedasticity test**

```{r}
bptest(women_final)
```

Homoscedacity assumption is not violated, p \> 0.05

**Independence of Errors test:**

```{r}
durbinWatsonTest(women_final)
```

Independence of errors assumption is not violated p \< 0.05.

**Conclusion:** The model show good fit.

#### Polinomynal regression analysis for testosterone

```{r}
quadratic_model_women <- lm(accu  ~ poly(Tes, 2, raw = TRUE), data = women_data)
summary(quadratic_model_women)
```

No quadratic effect of testosterone.

### 1.3.2. Men

**Tidying the table**

```{r}
#| echo: false
men_data <- CST_full_parameters |> filter(Sex == 2) |> drop_na(EA1, EA2, Tiredness_1, Tiredness_2)
```

#### Running first LM model with all the variables

```{r}
men_lm_model <- lm(accu~ Age + STEM + PANAS_PA + PANAS_NA + EA1 + EA2 + Tiredness_1 + Tiredness_2 + Femininity + Masculinity +Tes, data = men_data)
```

```{r}
summary(men_lm_model)
```

#### AIC procedure

```{r}
stepAIC(men_lm_model, direction = "backward")
```

#### Final model

```{r}
men_final <- lm(accu ~ EA1, data = men_data)
```

```{r}
summary(men_final)
```

#### Assumptions testing

```{r}
#| echo: false
par(mfrow = c(2, 2))
plot(men_final)
```

**Normality of residuals**

```{r}
#| echo: false
#| output: false
model.metrics.men <- augment(men_final) |> 
  dplyr::select(-.hat, -.sigma, -.fitted)

shapiro_test(model.metrics.men$.resid)
```

```{r}
ad.test(model.metrics.men$.resid)
```

IN QQ plot we can see that there is three outliers, otherwise the plot looks normal. At larger sample sizes the Shapiro-Wilk test becomes very sensitive even to a minor deviation from normality. Anderson-Darling normality test p \> 0.05, therefore I assume normal distribution.

**Homoscedasticity test**

```{r}
bptest(men_final)
```

Probably 17, 24, 30 are "outliers" which are effecting homoscedacity. Homoscedacity assuption is violated

**Independence of Errors test:**

```{r}
durbinWatsonTest(men_final)
```

The independence of error assumption is violated.

**Conclusion:** Some assumptions are violated therefore the model should be interpreted with cautious.

#### Polinomynal regression analysis for testosterone

```{r}
quadratic_model_men <- lm(accu ~ poly(Tes, 2, raw = TRUE), data = men_data)
```

```{r}
summary(quadratic_model_men)
```

No quadratic effect of testosterone

### 1.4 Mediation analysis

*Output of mediation analysis is not accessible as PROCESS v4.3 functions available as script, not a R package.*

```{r}
#| eval: false
process (data = CST_full_parameters, y = "accu", x = "Sex", m ="EA2", model = 4, effsize =1, total =1, stand =1, cov = "STEM", boot = 10000 , seed = 654321)
```
Percentage of the sex difference explained by EA2 


```{r}
Proc_EA <- (0.1303*100)/0.6075
Proc_EA
```

